The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-03 analysis/paper/paper.qmd
Document word embeddings script

# Load packages ----
## List packages to be loaded
packages <- c(
  "here",                                        # Managing paths
  "tidyverse", "reshape2",                       # Data manipulation and visualisation
  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
  "tidygeocoder",                                # Working with spatial data
  "igraph", "ggraph"                             # Network analysis and visualization
)

@@

## Add lemmas as optional column to ngram function

## Bigrams
emu_theses_bigrams <- emu_theses |> 
  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)

spacyr::spacy_parse(emu_theses_bigrams$first)

@@

bigram_stem_counts <- emu_theses_bigrams |> 
  count(first_stem, second_stem, sort = TRUE)

bigram_counts
bigram_stem_counts

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text")

trigram_counts <- emu_theses_trigrams |> 
  count(first, second, third, sort = TRUE)

@@

## Bigram tf-idf
bigram_tf_idf <- emu_theses_bigrams |> 
  count(title, ngram) |> 
  bind_tf_idf(ngram, title, n) |> 
  arrange(desc(tf_idf))

@@

## Trigram tf-idf
trigram_tf_idf <- emu_theses_trigrams |> 
  count(title, ngram) |> 
  bind_tf_idf(ngram, title, n) |> 
  arrange(desc(tf_idf))

@@

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
  geom_node_point(color = "lightblue", linewidth = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

@@

emu_theses_bigrams_top <- emu_theses_bigrams |>
  group_by(title) |>
  count(ngram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() |>
  mutate(bigram = reorder(ngram, n))

emu_theses_bigrams_top_all <- emu_theses_bigrams |>
  count(ngram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  mutate(ngram = reorder(ngram, n))

emu_theses_bigrams_top |>
  filter(title == unique(title)[1:10]) |> 
  ggplot(aes(x = ngram, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

@@

The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words. As such, it shows words that are the most important to one document in a collection of documents [@silge2017].

>>>>

# Load packages ----
## List packages to be loaded
packages <- c(
  "here",                                        # Managing paths
  "tidyverse", "reshape2", "widyr",              # Data manipulation and visualisation
  "furrr",                                       # Parallel use of purrr::map() functions
  "tidytext", "SnowballC", "tm", "topicmodels",  # Text processing and analysis
  "tidygeocoder",                                # Working with spatial data
  "igraph", "ggraph"                             # Network analysis and visualization
)

@@

## TO DO: generalise bigrams and trigrams into ngrams function

## Bigrams
emu_theses_bigrams <- emu_theses |> 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
  anti_join(stop_words, by = c("first" = "word")) |> 
  anti_join(stop_words, by = c("second" = "word")) |> 
  filter(str_detect(first, "[a-z]") & 
         str_detect(second, "[a-z]"))

@@

bigram_counts

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
  filter(!first %in% stop_words$word,
         !second %in% stop_words$word,
         !third %in% stop_words$word) |>  
  filter(str_detect(first, "[a-z]"), 
         str_detect(second, "[a-z]"),
         str_detect(third, "[a-z]"))

trigram_counts <- emu_theses_trigrams |> 
  count(first, second, sort = TRUE)

@@

## Bigram tf-idf
bigram_tf_idf <- emu_theses_bigrams |> 
  count(title, bigram) |> 
  bind_tf_idf(bigram, title, n) |> 
  arrange(desc(tf_idf))

@@

## Trigram tf-idf
trigram_tf_idf <- emu_theses_trigrams |> 
  count(title, trigram) |> 
  bind_tf_idf(trigram, title, n) |> 
  arrange(desc(tf_idf))

@@

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

@@

emu_theses_bigrams_top <- emu_theses_bigrams |>
  group_by(title) |>
  count(bigram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() |>
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top_all <- emu_theses_bigrams |>
  count(bigram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  mutate(bigram = reorder(bigram, n))

emu_theses_bigrams_top |>
  filter(title == unique(title)[1:6]) |> 
  ggplot(aes(x = bigram, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

@@

Describe word embeddings ...

```{r word-embeddings}
#| include: false

# Prepare tidy data frame for word embeddings
tidy_emu_theses <- emu_theses |> 
  select(title, text_clean) |> 
  unnest_tokens(word, text_clean) |> 
  add_count(word) |> 
  filter(n >= 50) |> 
  select(-n)

# Create a nested data frame with one row per thesis
nested_words <- tidy_emu_theses |> 
  nest(words = c(word))

```{r word-embeddings}

# Create sliding window function to calculate skipgram probabilities,
# that is, how often each word occurs near each other word.
# A small window size of 3-4 shows how a word is used, while a larger  
# window size of, say, 10 gives more information about the domain topic.
slide_windows <- function(tbl, window_size) {
  # Iterate through an input vector with a fixed window size
  skipgrams <- slider::slide(
    tbl,
    ~.x,
    .after = window_size - 1,
    .step = 1L,
    .complete = TRUE  # Only evaluate function on complete windows
  )

  # Run even when unsuccessful
  safe_mutate <- safely(mutate)

  # Map safe_mutate to skipgrams 
  out <- map2(skipgrams,
              1:length(skipgrams),
              ~ safe_mutate(.x, window_id = .y))

  out |> 
    transpose() |>  # Transpose `result` and `error` into columns,
    pluck("result") |>  # only keep `result`,
    compact() |>  # remove any NULL values [is this for the last windows?],
    bind_rows()  # and bind all rows into a single data frame.
}
```

```{r word-embeddings}

# To reduce computation time, we use the {furrr} 
# and underlying {future} packages for parallel processing
future::plan(multisession)

# To calculate how often words occur on their own or tigether with other words
# we use point-wise mutual information (PMI)
tidy_pmi <- nested_words |> 
  mutate(words = future_map(words, slide_windows, 4L)) |> 
  unnest(words) |> 
  unite(window_id, title, window_id) |> 
  pairwise_pmi(word, window_id)

The high-dimensional sparse matrix of word features is projected into reduced, 100-dimensional set of features.

```{r word-embeddings}

# Determine word vectors using SVD, a method for dimensionality reduction
# The resulting matrix will be of size n_voc * n_dim instead of n_voc * n_voc
tidy_word_vectors <- tidy_pmi |> 
  widely_svd(
    item1, item2, pmi,
    nv = 100, maxit = 1000
  )

```{r word-embeddings}

# Calculate nearest neighbours of a given word
nearest_neighbors <- function(df, token) {
  df |> 
    widely(
      ~ {
        y <- .[rep(token, nrow(.)), ]
        res <- rowSums(. * y) /
          (sqrt(rowSums(. ^ 2)) * sqrt(sum(.[token, ] ^ 2)))

        matrix(res, ncol = 1, dimnames = list(x = names(res)))
      },
      sort = TRUE
    )(item1, dimension, value) |> 
    select(-item2)
}

tidy_word_vectors |> 
  nearest_neighbors("price")
```

```{r word-embeddings}

# The first n components with top words found in the corpus of theses
tidy_word_vectors |> 
  filter(dimension <= 24) |> 
  group_by(dimension) |> 
  top_n(12, abs(value)) |> 
  ungroup() |> 
  mutate(item1 = reorder_within(item1, value, dimension)) |> 
  ggplot(aes(item1, value, fill = dimension)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~dimension, scales = "free_y", ncol = 4) +
  scale_x_reordered() +
  coord_flip() +
  labs(
    x = NULL,
    y = "Value",
    title = "First 24 principal components for text of theses",
    subtitle = "Top words contributing to the components that explain the most variation"
  )

```{r word-embeddings}

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_emu_theses |> 
  count(title, word) |> 
  cast_sparse(title, word, n)

embedding_matrix <- tidy_word_vectors |> 
  cast_sparse(item1, dimension, value)

doc_matrix <- word_matrix %*% embedding_matrix

dim(doc_matrix)

doc_matrix
